---
title: "Regression Project.Rmd"
author: "Harold Trammel"
date: "September 7, 2015"
output: pdf_document
---
```{r global_options, include=TRUE, echo=FALSE}
knitr::opts_chunk$set(fig.width=12, fig.height=7, fig.path='Figs/',
                      echo=TRUE, warning=FALSE, message=FALSE)
options(scipen=1, digits=2)
fig <- 0
tbl <- 0

data("mtcars")
require(memisc)
require(pander)
require(ggplot2)
require(dplyr)
require(knitr)

data("mtcars")

cars <- data.frame(mtcars)
```

## Motor Trend Car Road Tests

### Executive Summary

### Introduction
Motor Trend is interested in exploring the relationship between a set of variables and miles per gallon (MPG) (outcome). They have provided a set of data from automobiles that they tested over a period of time.  They are particularly interested in the following two questions:

* “Is an automatic or manual transmission better for MPG”
* "Quantify the MPG difference between automatic and manual transmissions"

A modern automobile is a complex machine and   would suggest that no one variable is fully responsible for its performance and efficiency


***
A data frame with 32 observations on 11 variables.
```
[, 1]	mpg	    Miles/(US) gallon
[, 2]	cyl	    Number of cylinders
[, 3]	disp	Displacement (cu.in.)
[, 4]	hp	    Gross horsepower
[, 5]	drat	Rear axle ratio
[, 6]	wt	    Weight (lb/1000)
[, 7]	qsec	1/4 mile time
[, 8]	vs	    V/S
[, 9]	am	    Transmission (0 = automatic, 1 = manual)
[,10]	gear	Number of forward gears
[,11]	carb	Number of carburetors
```
***

```{r pander_attempt}
lm0 <- lm(hp ~ wt, mtcars)
lm1 <- lm(qsec ~ hp, mtcars)
lm2 <- lm(qsec ~ wt, mtcars)
mtable123 <- mtable('Model 1' = lm0,
            'Model 2' = lm1,
            'Model 3' = lm2,
            summary.stats = c('R-squared','F','p','N'))
pander(mtable123)
```


## Fit all variables against MPG
```{r all_fit}
all_fit <- lm(mpg ~ . , data = cars)
all_fit_coef <- summary(all_fit)$coef
kable(all_fit_coef, caption="All Fit Coefficients")
```

## Fit automatic vs manual against MPG
```{r am_fit}
# [, 9]	am	    Transmission (0 = automatic, 1 = manual)
am_fit <- lm(mpg ~ am, data = cars)
am_fit_coef <- summary(am_fit)$coef
kable(am_fit_coef, caption="All Fit Coefficients")

g_am = ggplot(cars, aes(x = am, y = mpg)) + 
    ylab("Miles per gallon") + 
    xlab("Automatic or Manual") + 
    geom_point(aes(y=mpg), size = 7, colour = "black", alpha=0.5) + 
    geom_point(aes(x = am),size = 5, colour = "blue", alpha=0.6) +
    geom_smooth(method = "lm", colour = "black") 
print(g_am)

```

## Fit Weight against MPG
```{r wt_fit}
# [, 6]	wt	    Weight (lb/1000)
wt_fit <- lm(mpg ~ wt, data = cars)
wt_fit_coef <- summary(wt_fit)$coef
kable(wt_fit_coef, caption="Weight Coefficients")

g_wt = ggplot(cars, aes(x = wt, y = mpg)) + 
    ylab("Miles per gallon") + 
    xlab("Weight lb/1000") + 
    geom_point(aes(y=mpg), size = 7, colour = "black", alpha=0.5) + 
    geom_point(aes(x = wt),size = 5, colour = "red", alpha=0.6) +
    geom_smooth(method = "lm", colour = "black") 
print(g_wt)
```

## Fit engine displacement against MPG
```{r disp_fit}
# [, 3]	disp	Displacement (cu.in.)
disp_fit <- lm(mpg ~ am + gear, data = cars)
disp_fit_coef <- summary(disp_fit)$coef
kable(disp_fit_coef, caption="Displacement Coefficients")

g_disp = ggplot(cars, aes(x = disp, y = mpg)) + 
    ylab("Miles per gallon") + 
    xlab("Engine Displacement") + 
    geom_point(aes(y=mpg), size = 7, colour = "black", alpha=0.5) + 
    geom_point(aes(x = disp),size = 5, colour = "yellow", alpha=0.6) +
    geom_smooth(method = "lm", colour = "black") 
print(g_disp)
```

## Fit Gross horsepower against MPG
```{r hp_fit}
# [, 4]	hp	    Gross horsepower
hp_fit <- lm(mpg ~ hp, data = cars)
hp_fit_coef <- summary(hp_fit)$coef
kable(hp_fit_coef, caption="Horsepower Coefficients")

g_hp = ggplot(cars, aes(x = hp, y = mpg)) + 
    ylab("Miles per gallon") + 
    xlab("Horsepower") + 
    geom_point(aes(y=mpg), size = 7, colour = "black", alpha=0.5) + 
    geom_point(aes(x = hp),size = 5, colour = "green", alpha=0.6) +
    geom_smooth(method = "lm", colour = "black") 
print(g_hp)
```

## Fit number of gears against MPG
```{r gear_fit}
# [,10]	gear	Number of forward gears
gr_fit <- lm(mpg ~ gear, data = cars)
gr_fit_coef <- summary(gr_fit)$coef
kable(gr_fit_coef, caption="Numb. Gears Coefficients")

g_gr = ggplot(cars, aes(x = gear, y =mpg)) + 
    ylab("Miles per gallon") + 
    xlab("Numb. Gears") + 
    geom_point(aes(y=mpg), size = 7, colour = "black", alpha=0.5) + 
    geom_point(aes(x = gear),size = 5, colour = "orange", alpha=0.6) +
    geom_smooth(method = "lm", colour = "black") 
print(g_gr)
```

## Fit number of carburators against MPG
```{r carb_fit}
# [,11]	carb	Number of carburetors
carb_fit <- lm(mpg ~ carb, data = cars)
carb_fit_coef <- summary(carb_fit)$coef
kable(carb_fit_coef, caption="Numb. Carburator Coefficients")

g_carb = ggplot(cars, aes(x = carb, y =mpg)) + 
    ylab("Miles per gallon") + 
    xlab("Numb. Carburators") + 
    geom_point(aes(y=mpg), size = 7, colour = "black", alpha=0.5) + 
    geom_point(aes(x = carb),size = 5, colour = "purple", alpha=0.6) +
    geom_smooth(method = "lm", colour = "black") 
print(g_carb)
```

## Fit Rear Gear Ratio  against MPG
```{r drat_fit}
# [, 5]	drat	Rear axle ratio
drat_fit <- lm(mpg ~ drat, data = cars)
drat_fit_coef <- summary(drat_fit)$coef
kable(drat_fit_coef, caption="Rear Gear Ratio Coefficients")

g_drat = ggplot(cars, aes(x = drat, y =mpg)) + 
    ylab("Miles per gallon") + 
    xlab("Rear Axle Ratio") + 
    geom_point(aes(y=mpg), size = 7, colour = "black", alpha=0.5) + 
    geom_point(aes(x = drat),size = 5, colour = "pink", alpha=0.6) +
    geom_smooth(method = "lm", colour = "black") 
print(g_drat)
```

### Fit Vee vs Straight against MPG
```{r vs_fit}
# [, 8]	vs	    V/S
vs_fit <- lm(mpg ~ vs, data = cars)
vs_fit_coef <- summary(vs_fit)$coef
kable(vs_fit_coef, caption="Vee vs Straight Coefficients")

g_vs = ggplot(cars, aes(x = vs, y =mpg)) + 
    ylab("Miles per gallon") + 
    xlab("Vee or Straight") + 
    geom_point(aes(y=mpg), size = 7, colour = "black", alpha=0.5) + 
    geom_point(aes(x = vs),size = 5, colour = "cyan", alpha=0.6) +
    geom_smooth(method = "lm", colour = "black") 
print(g_vs)
```

## Fit number of cylinders against MPG
```{r cyl_fit}
# [, 2]	cyl	    Number of cylinders
cyl_fit <- lm(mpg ~ cyl, data = cars)
cyl_fit_coef <- summary(cyl_fit)$coef
kable(cyl_fit_coef, caption="Numb. Cylinders Coefficients")

g_cyl = ggplot(cars, aes(x = cyl, y =mpg)) + 
    ylab("Miles per gallon") + 
    xlab("Numb. Cylinders") + 
    geom_point(aes(y=mpg), size = 7, colour = "black", alpha=0.5) + 
    geom_point(aes(x = cyl),size = 5, colour = "magenta", alpha=0.6) +
    geom_smooth(method = "lm", colour = "black") 
print(g_cyl)
```

## Fit Quarter Mile Time against MPG
```{r qsec_fit}
# [, 7]	qsec	1/4 mile time
qsec_fit <- lm(mpg ~ qsec, data = cars)
qsec_fit_coef <- summary(qsec_fit)$coef
kable(qsec_fit_coef, caption="Quarter Mile Time Coefficients")

qsec_cyl = ggplot(cars, aes(x = qsec, y =mpg)) + 
    ylab("Miles per gallon") + 
    xlab("Quarter Mile Time (sec)") + 
    geom_point(aes(y=mpg), size = 7, colour = "black", alpha=0.5) + 
    geom_point(aes(x = qsec),size = 5, colour = "tan", alpha=0.6) +
    geom_smooth(method = "lm", colour = "black") 
print(qsec_cyl)
```